---
title: "Final Presentation"
author: "Team Jacob"
date: "04/21/2025"
format: revealjs
---

# Goal

-   Predict **Profit** of new CravenSpeed products\
-   Evaluate models using **RMSE**
-   Use exactly **10 engineered features**\
-   Train models using **caret** with 5-fold cross-validation

------------------------------------------------------------------------

# Feature Engineering Strategy

We created 10 features based on product cost, sales, and design:

| Feature                       | Rationale                          |
|-------------------------------|------------------------------------|
| Unit Margin                   | Retail price - BOM cost            |
| Log Units Sold                | Handles skewed sales               |
| Product Age                   | Older products may stagnate        |
| Is Craven                     | Distinguishes internal vs external |
| Material Type                 | Reflects cost/quality              |
| MFG Process                   | Indicates outsourcing vs in-house  |
| Weight/Price Ratio            | Efficiency or premium              |
| \% Direct Sales               | Profit often higher direct         |
| Sales Channel                 | Some may have higher margins       |
| Interaction: Units × % Direct | Captures sales style impact        |

------------------------------------------------------------------------

# Engineer Function

```{r}
engineer <- function(df) {
  df <- df %>%
    mutate(
      unit_margin = retailPrice - `BOM Cost`,
      log_units_sold = log1p(`Units Sold`),
      product_age = as.numeric(difftime(Sys.Date(), 
                          suppressWarnings(as.Date(`First Order (from sales data)`, format = "%m/%d/%y")), 
                          units = "days")) / 365,
      is_craven = `Product Type` == "#Craven",
      material_type = as.factor(`Main Component Material`),
      mfg_process = as.factor(`Main Component MFG Process`),
      weight_price_ratio = `Unit Weight (pounds)` / (retailPrice + 1),
      pct_direct_sales = as.numeric(gsub("%", "", `pct_Direct Sales`)) / 100,
      sales_channel = as.factor(`Sales Channel`),
      direct_volume = `Units Sold` * pct_direct_sales
    ) %>%
    select(unit_margin, log_units_sold, product_age, is_craven, 
           material_type, mfg_process, weight_price_ratio, 
           pct_direct_sales, sales_channel, direct_volume)
  return(df)
}
```

------------------------------------------------------------------------

# Modeling Workflow

```{r}
library(tidyverse)
library(caret)

fast <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/refs/heads/master/dat/craven_train.rds")))

# Engineer features
profit <- fast$`Revenue 2019 to present` - fast$`BOM Cost` * fast$`Units Sold`
fast_features <- engineer(fast)
fast_features$Profit <- profit
fast_features <- na.omit(fast_features)

# Remove single-level factors
fast_features <- fast_features %>%
  select(where(~ !is.factor(.) || nlevels(.) > 1))

# Train control
ctrl <- trainControl(method = "cv", number = 5)
```

------------------------------------------------------------------------

# Linear Model (LM)

```{r}
model_lm <- train(Profit ~ ., data = fast_features, method = "lm", trControl = ctrl)
model_lm$results
```

------------------------------------------------------------------------

# Random Forest (RF)

```{r}
model_rf <- train(Profit ~ ., data = fast_features, method = "rf", trControl = ctrl, tuneLength = 3)
model_rf$results
```

------------------------------------------------------------------------

# RMSE Comparison

```{r}
results <- bind_rows(
  model_lm$results %>% mutate(Model = "Linear Regression"),
  model_rf$results %>% mutate(Model = paste0("Random Forest (mtry=", mtry, ")"))
)

ggplot(results, aes(x = Model, y = RMSE)) +
  geom_col(fill = "steelblue") +
  labs(title = "RMSE Comparison", y = "RMSE", x = "")
```

------------------------------------------------------------------------

# Feature Importance (Random Forest)

```{r}
plot(varImp(model_rf), top = 10)
```

------------------------------------------------------------------------

# Key Insights

-   Random Forest outperforms Linear Regression (lower RMSE)
-   Top features: `unit_margin`, `log_units_sold`, `direct_volume`
-   Products with high direct sales volume and strong unit margins yield better profits

------------------------------------------------------------------------

# Limitations

-   Some features had little variance (e.g., materials or sales channel)
-   Missing sales dates limited usable data
-   Profit is based only on revenue and BOM, not all costs

------------------------------------------------------------------------

# Next Steps

-   Try `xgboost` or `ridge` regression
-   Add more nuanced time-series or seasonal features
-   Incorporate other cost factors if available

------------------------------------------------------------------------

# Thank You!

Questions?